Optimal Pricing and Informal Sharing: \\ Evidence from Piped Water in Manila

### 2018-11-08 phil_paper/jmp_wjv_v3.tex
lots of paper edits

\usepackage[notablist,nofiglist]{endfloat}
%\renewcommand{\efloatseparator}{\mbox{}}

\renewcommand{\thetable}{\Roman{table}}
\renewcommand{\thefigure}{\Roman{figure}}

\renewcommand{\thesection}{\Roman{section}} 
\renewcommand{\thesubsection}{\thesection.\Alph{subsection}}
\renewcommand{\thesubsubsection}{\thesubsection.\roman{subsubsection}}

%\usepackage[dvipsnames]{xcolor}
%\usepackage{cite}
\onehalfspacing
\usepackage{natbib}  
%\title{ Optimal Pricing and Informal Sharing: \\ Evidence from Piped Water in Manila }
\title{ OPTIMAL PRICING AND INFORMAL SHARING: \\ EVIDENCE FROM PIPED WATER IN MANILA}

@@

\textbf{Sample} : The estimation uses a 10\% sample of connections from the connection survey (5,000 connections), which serves as the primary sample for the counterfactual exercises.  This sample is combined with 2,151 connections from the quasi-experiment examined in the second step of the estimation (Section~\ref{section:hasslecostestimation}), which takes preference estimates from this step as an input.  Together, these connections produce 349,265 months of consumption observations used in the preference estimation.  All standard errors are bootstrapped using \input{tables/boot_max}repetitions.  This sample does not include households using water from vendors since the data do not include information on consumption for these households.  Section~\ref{section:fixedcostestimation} details the imputation procedure used to infer preferences for these households.  
%Appendix~\ref{appendix:sigmanuassumptions} determines the ex-ante assumptions on $\sigma_{\epsilon}$ needed to identify the particular demand function.
%The estimates are robust to varying the total and relative sample sizes of these groups.

@@

Table~\ref{table:elasticity} presents connection-level demand elasticity estimates across terciles of (1) imputed income for the connection owner and (2) average monthly usage for the connection.  Elasticities exhibit little heterogeneity across imputed income levels, but substantial variation across usage levels with larger users appearing much less sensitive to prices than smaller users.  Since many low-income households share connections with neighbors, there is a weak correlation between income and usage-level, which helps account for the lack of heterogeneity in elasticity.\footnote{Noise in the income imputation process may also weaken any correlation between elasticity and income.}  Heterogeneity in elasticities according to monthly usage is consistent with moral hazard when households share a connection:  increasing the number of households per connection increases the water use for those connections while also exacerbating free-riding behavior.  Different water using activities across the water demand profile may also produce this heterogeneity in elasticities.  This heterogeneity provides a potential motivation for usage-based price discrimination.  For example, matching larger, less price-sensitive users with higher prices may allow for a more efficient allocation of water while still covering costs of production.
\begin{table}
\centering
\begin{threeparttable}
\caption{Demand Elasticities for Water Connections}\label{table:elasticity}
\input{tables/elasticity_estimates} 
\begin{tablenotes}
\item {\footnotesize Elasticities are calculated from 7,151 connections making 349,265 connection-months.  Income is imputed based on demographics, locations, and their interactions.  Standard errors are bootstrapped at the connection level.}
\end{tablenotes}
\end{threeparttable}
\end{table}
The demand estimation yields an average elasticity of \input{tables/mean_elasticity}\unskip, which can be interpreted as the percentage decrease in quantity due to a one percentage increase in price.  This estimate also closely matches reduced-form elasticity estimates from price-change quasi-experiment detailed in Appendix~\ref{appendix:ratechangeelasticity}.  This estimate is in range of the 0.98 estimate from \cite{szabo2015value} as well as similar studies in the developing world that find elasticities between 0.01 and 0.81 (\cite{diakite2009proposal}, \cite{strand2005water}). 
%is its inability to match zero consumption values observed in the data
\begin{table}
\caption{Model Performance}\label{table:modelperformance}
\input{tables/sample_prediction}
\end{table}
Table~\ref{table:modelperformance} compares predicted consumption from the model to observed consumption in the data.  This exercise finds negligible differences across average consumption and the variance of consumption for connections according to sharing status.  Closely matching observed patterns, this model improves on previous approaches to demand estimation likely by recovering a separate fixed preference term for each connection as well as rich heterogeneity in the variance terms (\cite{szabo2015value}).

>>>>

%\usepackage[notablist,nofiglist]{endfloat}
%\renewcommand{\efloatseparator}{\mbox{}}

%\renewcommand{\thetable}{\Roman{table}}
%\renewcommand{\thefigure}{\Roman{figure}}

%\renewcommand{\thesection}{\Roman{section}} 
%\renewcommand{\thesubsection}{\thesection.\Alph{subsection}}
%\renewcommand{\thesubsubsection}{\thesubsection.\roman{subsubsection}}

%\usepackage[dvipsnames]{xcolor}
%\usepackage{cite}
\onehalfspacing
\usepackage{natbib}  
%\title{ Optimal Pricing and Informal Sharing: \\ Evidence from Piped Water in Manila }
\title{ Optimal Pricing and Informal Sharing: \\ Evidence from Piped Water in Manila }

@@

\textbf{Sample} : The estimation uses a 10\% sample of connections from the connection survey (5,000 connections), which serves as the primary sample for the counterfactual exercises.  This sample is combined with 2,151 connections from the quasi-experiment examined in the second step of the estimation (Section~\ref{section:hasslecostestimation}), which takes preference estimates from this step as an input.  Together, these connections produce 349,265 months of consumption observations used in the preference estimation.  All standard errors are bootstrapped using \input{tables/boot_max}repetitions.  This sample does not include households using water from vendors since the data do not include information on consumption for these households.  Section~\ref{section:fixedcostestimation} details the imputation procedure used to infer preferences for these households.  Computationally, the estimation minimizes the likelihood function using a trust-region algorithm provided with the analytical gradient vector and Hessian matrix of the likelihood function.
%Appendix~\ref{appendix:sigmanuassumptions} determines the ex-ante assumptions on $\sigma_{\epsilon}$ needed to identify the particular demand function.
%The estimates are robust to varying the total and relative sample sizes of these groups.

@@

Table~\ref{table:elasticity} presents connection-level demand elasticity estimates across terciles of (1) imputed income for the connection owner and (2) average monthly usage for the connection.  Elasticities exhibit little heterogeneity across imputed income levels, but substantial variation across usage levels with larger users appearing much less sensitive to prices than smaller users.  Since many low-income households share connections with neighbors, there is a weak correlation between income and usage-level, which helps account for the lack of heterogeneity in elasticity.\footnote{Noise in the income imputation process may also weaken any correlation between elasticity and income.}  Heterogeneity in elasticities according to monthly usage is consistent with moral hazard when households share a connection:  increasing the number of households per connection increases the water use for those connections while also exacerbating free-riding behavior.  Different water using activities across the water demand profile may also produce this heterogeneity in elasticities.  This heterogeneity provides a potential motivation for usage-based price discrimination.  For example, matching larger, less price-sensitive users with higher prices may allow for a more efficient allocation of water while still covering costs of production.
\begin{table}
\centering
\begin{threeparttable}
\caption{Demand Elasticities for Water Connections}\label{table:elasticity}
\input{tables/elasticity_estimates} 
\begin{tablenotes}
\item {\footnotesize Elasticities are calculated from 7,151 connections making 349,265 connection-months.  Income is imputed based on demographics, locations, and their interactions.  Standard errors are bootstrapped at the connection level.}
\end{tablenotes}
\end{threeparttable}
\end{table}
The demand estimation yields an average elasticity of \input{tables/mean_elasticity}\unskip, which can be interpreted as the percentage decrease in quantity due to a one percentage increase in price.  This estimate also closely matches reduced-form elasticity estimates from price-change quasi-experiment detailed in Appendix~\ref{appendix:ratechangeelasticity}.  This estimate is in range of the 0.98 estimate from \cite{szabo2015value} as well as similar studies in the developing world that find elasticities between 0.01 and 0.81 (\cite{diakite2009proposal}, \cite{strand2005water}). 
%is its inability to match zero consumption values observed in the data
\begin{table}
\centering
\caption{Model Performance}\label{table:modelperformance}
\input{tables/sample_prediction}
\end{table}
Table~\ref{table:modelperformance} compares predicted consumption from the model to observed consumption in the data.  This exercise finds negligible differences across average consumption and the variance of consumption for connections according to sharing status.  Closely matching observed patterns, this model improves on previous approaches to demand estimation likely by recovering a separate fixed preference term for each connection as well as rich heterogeneity in the variance terms (\cite{szabo2015value}).
